SU($n$) spin models near the Heisenberg critical point

### 2020-03-25 collective_sun/sun_perturbation.tex
fix section on the multi-body eigenvalue problem to account for multi-body operators without permutational symmetry some additional minor fixes elsewhere

\newcommand{\floor}[1]{\lfloor{#1}\rfloor}
\newcommand{\ceil}[1]{\lceil{#1}\rceil}

@@

\usepackage{mathtools} % for coloneqq

@@

\red{[TODO: fix section to consider the case without permutational
  symmetry of $X_{\m w}$]}

In this section, we simplify the products of the form
$H_0X_{\m w}\ket\psi$ to arrive at the expansion in
\eqref{eq:diagnosis}, which defines conditions under which a
multi-body perturbation $X_{\m w_\Delta}$ generates eigenstates of the
SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when
acting on a permutationally symmetric state $\ket\psi\in\M_0$.  Here
$X_{\m w}$ is an $M$-body perturbation defined by
\begin{align}
  X_{\m w}
  \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k
  = \f1{M!} \sum_{\sigma\in\pi\p{M}} \sum_{k\in\C_N\p{M}}
  w_{\sigma\p{k}} X_{\sigma\p{k}},
\end{align}
where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
a list of $M$ spins that $X_k$ acts on; $\m w$ is a dimension-$M$
(i.e.~$M$-index) tensor with scalar entries $w_k$; $\D_N\p{M}$ is the
strictly off-diagonal part of $\ZZ_N^M$, defined in
\eqref{eq:off_diags}; $\pi\p{M}$ is the permutation group of $\ZZ_M$;
and $\C_N\p{M}$ is the set of subsets (``choices'') of $M$ elements
from $\ZZ_N$, defined in \eqref{eq:choices}.

@@

In order to build familiarity with the problem at hand, we first
consider the simple case of a single-body perturbation and simplify
\begin{align}
  H_0 X_{\m w} \ket\psi
  = \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\ZZ_N}}
  h_{pq} w_k \Pi_{pq} X_k \ket\psi.
  \label{eq:single_body_diagnosis_start}
\end{align}
The sum in \eqref{eq:single_body_diagnosis_start} has terms with
$k\in\set{p,q}$, and terms with $k\notin\set{p,q}$.  In the case of
$k\notin\set{p,q}$, the permutation operator $\Pi_{pq}$ commutes with
$X_k$ and annihilates on $\ket\psi$, leaving terms at each fixed $k$
of the form
\begin{align}
  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\notin\set{p,q}}} h_{pq}
  = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}
  - \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\set{p,q}}} h_{pq}
  = E_0 - \f12 \sum_{\substack{p,q\in\ZZ_N\\k\in\set{p,q}}} h_{pq}
  = E_0 - h_k,
\end{align}
where we define $h_{pq}$ for arbitrary $p,q$ by enforcing
$h_{pq}=h_{qp}$ and $h_{pp}=0$; and $h_k \equiv \sum_\ell h_{k\ell}$.
The terms in \eqref{eq:single_body_diagnosis_start} with
$k\notin\set{p,q}$ are then
\begin{align}
  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\notin\set{p,q}}}
  h_{pq} w_k \Pi_{pq} X_k \ket\psi
  = \sum_{k\in\ZZ_N} \p{E_0 - h_k} w_k X_k \ket\psi
  = E_0 X_{\m w} \ket\psi - \sum_{k\in\ZZ_N} h_k w_k X_k \ket\psi
\end{align}
while the terms in \eqref{eq:single_body_diagnosis_start} with
$k\in\set{p,q}$ take the form
\begin{align}
  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\set{p,q}}}
  h_{pq} v_k \Pi_{pq} X_k \ket\psi
  = \sum_{\p{k,q}\in\C_N\p{2}} h_{kq} w_k X_q \ket\psi
  + \sum_{\p{p,k}\in\C_N\p{2}} h_{pk} w_k X_p \ket\psi
  = \sum_{k,q\in\ZZ_N} h_{kq} w_k X_q \ket\psi
\end{align}
which implies
\begin{align}
  H_0 X_{\m w} \ket\psi
  = E_0 X_{\m w} \ket\psi
  + \sum_{p\in\ZZ_N} \p{\sum_{q\in\ZZ_N} h_{pq} w_q - h_p w_p}
  X_p \ket\psi.
  \label{eq:single_body_diagnosis_end}
\end{align}
The action of the single-body perturbation $X_{\m w}$ on a
permutationally symmetric state therefore generates an eigenstate of
$H_0$ with energy $E_0+\Delta$ if the vector
$\v{\m w}\equiv\sum_{k\in\ZZ_N} w_k \ket{k}$ satisfies the eigenvalue
equation
\begin{align}
  \p{\m h - \diag\v h} \c \v{\m w} = \Delta \v{\m w},
  \label{eq:single_body_eig}
\end{align}
where $\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ is a matrix of
all couplings $h_{pq}$; the vector
$\v h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \ket{p}$ is the sum of all
columns of $\m h$; and $\diag\v h$ is a matrix with $\v h$ on the
diagonal and zeroes everywhere else.

When the interaction Hamiltonian $H_0$ is translationally invariant,
the single-body eigenvalue problem in \eqref{eq:single_body_eig} is
solvable analytically.  In this case, the couplings $h_{pq}$ depend
only on the separation $\abs{p-q}$, so the eigenvectors of $\m h$ are
plane waves of the form
\begin{align}
  \v{\m w}_k \equiv \sum_{p\in\ZZ_L^D} e^{ip\c k} \ket{p},
\end{align}
where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
sites are indexed by vectors $p\in\ZZ_L^D$, and wavenumbers take on
values $k\in\ZZ_L^D\times2\pi/L$.  The corresponding eigenvalues of
$\m h$ can be determined by expanding
\begin{align}
  \m h\c{\m w}_k
  = \sum_{p,q\in\ZZ_L^D} h_{pq} e^{iq\c k} \ket{p}
  = \sum_{p,d\in\ZZ_L^D} h_{p,p+d} e^{i\p{p+d}\c k} \ket{p}
  = \sum_{d\in\ZZ_L^D} h_{0,d} e^{id\c k} \v{\m w}_k
  = \sum_{d\in\ZZ_L^D} h_{0,d} \cos\p{d\c k} \v{\m w}_k,
\end{align}
where the imaginary terms vanish because $h_{0,d}=h_{0,-d}$.  The
remainder of \eqref{eq:single_body_eig} that we need to sort out is
$\diag\v h$, where all
$h_p=\sum_{q\in\ZZ_L^D}h_{pq}=\sum_{d\in\ZZ_L^D}h_{0,d}$ are equal,
which implies that $\diag\v h=\sum_{d\in\ZZ_L^D}h_{0,d}$ is a scalar.
We thus find that
\begin{align}
  \p{\m h - \diag\v h}\c\v v_k = \Delta_k \v{\m w}_k,
  &&
  \Delta_k \equiv \sum_{d\in\ZZ_L^D} h_{0,d} \sp{\cos\p{d\c k}-1}.
\end{align}

@@

\red{[TODO: fix section to consider the case without permutational
  symmetry of $X_{\m w}$]}

@@

  = \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\C_N\p{M}}}

@@

The sum in \eqref{eq:diagnosis_start} has terms with $p,q\in k$, terms
with $p,q\notin k$, and terms with only one of $p$ or $q\in k$.  The
permutation operator $\Pi_{pq}$ acts trivially on $X_k\ket\psi$ when
$p,q\in k$ or $p,q\notin k$, so

@@

  \sum_{\substack{k\in\C_N\p{M}\\\p{p,q}\in\C_N\p{2}\\
      p,q\in k~\t{or}~p,q\notin k}}

@@

  = \sum_{k\in\C_N\p{M}}
  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}}
  h_{pq} w_k X_k \ket\psi.

@@

For any fixed $k\in\C_N\p{M}$, we can simplify

@@

  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
  = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}
  - \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq}
  = E_0 - \tilde h_k,

@@

  \tilde h_k \equiv \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq},

@@

where we split a sum over $\p{p,q}\in\C_N\p{2}$ with the restriction
that $p,q\in k$ or $p,q\notin k$ into an unrestricted sum minus a
remainder with the complementary restriction; and $E_0$ is the
interaction energy of PS states $\ket\psi\in\M_0$.  The terms in
\eqref{eq:diagnosis_start} with $p,q\in k$ or $p,q\notin k$ are then

@@

  \sum_{k\in\C_N\p{M}} \p{E_0 - \tilde h_k} w_k X_k \ket\psi
  = E_0 X_{\m w} \ket\psi
  - \sum_{k\in\C_N\p{M}} \p{\tilde{\m h} * \m w}_k X_k \ket\psi,

@@

where $\tilde{\m h}$ is a tensor with entries $\tilde h_k$; and
$\tilde{\m h} * \m w$ denotes an element-wise (Kronecker) product of
$\tilde{\m h}$ and $\m w$, with entries $\tilde h_k w_k$.  The
remaining terms in \eqref{eq:diagnosis_start} with only one of

@@

  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \sum_{a\in\ZZ_M} h_{p k_a} w_k \Pi_{p k_a} X_k \ket\psi.

@@

where $k=\p{k_1,k_2,\cdots,k_M}$, with $k_a$ the $a$-th entry in $k$.
Defining $k_{a:p}$ to be a list that is equal to $k$ but with the
$a$-th number $k_a$ replaced by $p$,
$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$, we can
simplify

@@

  \Pi_{p k_a} X_k \ket\psi
  = \Pi_{p k_a} X_k \Pi_{p k_a} \ket\psi
  = X_{k_{a:p}}\ket\psi,
and switch the labels for $p$ and $k_a$ to find that
  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \sum_{a\in\ZZ_M} h_{pk_a} w_{k_{a:p}} X_k \ket\psi.
For each fixed $a$, the sum over $p$ above is essentially a
contraction of the matrix $\m h$ with the $a$-th index of the tensor
$\m w$.  Denoting such a contraction by $\m h\c_a\m w$, we can
therefore write this result as
  \sum_{k\in\C_N\p{M}} \sum_{a\in\ZZ_M}
  \p{\m h \circ_a \m w}_k X_k \ket\psi,
where, for clarity, the matrix elements of $\m h\circ\m w$ are
  \p{\m h \circ_a \m w}_k

@@

  h_{pk_a} w_{k_1 k_2 \cdots k_{a-1} p k_{a+1} \cdots k_M}.

@@

Altogether, we thus find that

@@

  = E_0 X_{\m w} \ket\psi + \sum_{k\in\C_N\p{M}}
  \hat{\m h}\p{\m w}_k X_k \ket\psi,

@@

  \equiv \sum_{a\in\ZZ_M} \m h \circ_a \m w - \tilde{\m h} * \m w.
  \label{eq:diagnosis_end}

@@

The action of the $M$-body perturbation $X_{\m w}$ on a
permutationally symmetric state therefore generates an eigenstate of
$H_0$ with energy $E_0+\Delta$ if the coefficient tensor $\m w$
satisfies the generalized eigenvalue equation
$\hat{\m h}\p{\m w} = \Delta \m w$, which can be written in the form

@@

@@

  \hat{\m h} \equiv \sum_{k\in\C_N\p{M}}
  \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \sum_{a\in\ZZ_M} h_{pk_a} \op{k}{k_{a:p}}
  - \sum_{k\in\C_N\p{M}} \tilde h_k \op{k},
  \v{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k \ket{k}.

@@

If $M=1$, then
$\p{\m h\circ\m w}_k=\sum_{\ell\in\ZZ_N} h_{k\ell} w_\ell$ and
$\tilde h_k = h_k \equiv \sum_{\ell\in\ZZ_N} h_{k\ell}$, so

@@

  = E_0 X_{\m w} \ket\psi + \sum_{k\in\ZZ_N}
  \p{\sum_{\ell\in\ZZ_N} h_{k\ell} w_\ell - h_k w_k} X_k \ket\psi,

@@

which is precisely what we found in
\eqref{eq:single_body_diagnosis_end}.  If $M=2$, then
$\p{\m h\circ\m w}_{pq} = \sum_{k\in\ZZ_N}
\p{h_{pk}w_{kq}+h_{qk}w_{pk}}$, so

@@

  = E_0 X_{\m w} \ket\psi + \sum_{\p{p,q}\in\C_N\p{2}}
  \sp{\sum_{k\in\ZZ_N}\p{h_{pk} w_{kq} + h_{qk} w_{pk}}
    - \tilde h_{pq} w_{pq}}
  X_k \ket\psi,

@@

where $\tilde h_{pq}=h_p+h_q-2h_{pq}$.

@@

Solving the eigenvalue problem in \eqref{eq:multi_body_eig_mat} to
find $M$-body perturbations $X_{\m w}$ that generate states of
definite interaction energy nominally requires diagonalizing the
matrix $\hat{\m h}$ with dimensions
${N\choose M}\times{N\choose M}\sim N^M\times N^M$.  If the
interaction Hamiltonian $H_0$ has any spatial symmetries, however,
then we can use these symmetries to reduce the complexity of this
eigenvalue problem.  Translational invariance, for example, can reduce
this problem to that of diagonalizing a matrix with dimensions
$\le{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times N^{M-1}$,
and isotropy can further reduce the problem size by factors that
depend on the lattice geometry.  The complexity of diagonalizing a
matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a reduction
of $d$ by a factor of 2 yields practically significant gains (a factor
of $2^3=8$) in the runtime of diagonalization.  Here, we provide a
general prescription for exploiting the spatial symmetries of $H_0$ to
simplify the eigenvalue problem in \eqref{eq:multi_body_eig_mat}.

Our task is essentially to express the multi-body eigenvalue problem
in \eqref{eq:multi_body_eig_mat} in a basis that respects the spatial
symmetries of $H_0$.  These symmetries essentially break the set of
choices $\C_N\p{M}$ into equivalence classes $\S_M$\footnote{For
  brevity, we will suppress the explicit dependence of $\S_M$ on $N$,
  as well as the dependence of $\S_M$ on any other factors such as
  lattice geometry.}, where elements $s\in\S_M$ are subsets
$s\subset\C_N\p{M}$ that are invariant under the appropriate symmetry
group action.  In the case of a translationally invariant and
isotropic system with two-body perturbations, for example, each
equivalence class $s\in\S_2$ would consist of all pairs
$\p{p,q}\in\C_N\p{2}$ that are a fixed distance $\abs{p-q}=d_s$ apart,
i.e.~$s=\set{\p{p,q}\in\C_N\p{2}:\abs{p-q}=d_s}$.

>>>>

\newcommand{\lra}{\leftrightarrow}
\newcommand{\floor}[1]{\lfloor{#1}\rfloor}

@@

@@

In this section, we simplify the products of the form
$H_0X_{\m w}\ket\psi$ to arrive at the expansion in
\eqref{eq:diagnosis}, which defines conditions under which a
multi-body perturbation $X_{\m w_\Delta}$ generates eigenstates of the
SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when
acting on a permutationally symmetric state $\ket\psi\in\M_0$.  Here
$X_{\m w}$ is an $M$-body perturbation defined by
\begin{align}
  X_{\m w}
  \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k
  = \f1{M!} \sum_{\sigma\in\pi\p{M}} \sum_{k\in\C_N\p{M}}
  w_{\sigma\p{k}} X_{\sigma\p{k}},
\end{align}
where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
a list of $M$ spins that $X_k$ acts on; $\m w$ is a dimension-$M$
(i.e.~$M$-index) tensor with scalar entries $w_k$; $\D_N\p{M}$ is the
strictly off-diagonal part of $\ZZ_N^M$, defined in
\eqref{eq:off_diags}; $\pi\p{M}$ is the permutation group of $\ZZ_M$;
and $\C_N\p{M}$ is the set of subsets (``choices'') of $M$ elements
from $\ZZ_N$, defined in \eqref{eq:choices}.  For shorthand, we will
define $\tilde X_{\m w}\equiv M! X_{\m w}$.

@@

In order to build familiarity with the problem at hand, we first
consider the simple case of a single-body perturbation and simplify
\begin{align}
  H_0 \tilde X_{\m w} \ket\psi
  = \sum_{\substack{k\in\ZZ_N\\\set{p,q}\in\C_N\p{2}}}
  h_{pq} w_k \Pi_{pq} X_k \ket\psi.
  \label{eq:single_body_diagnosis_start}
\end{align}
The sum in \eqref{eq:single_body_diagnosis_start} has terms with
$k\notin\set{p,q}$, and terms with $k\in\set{p,q}$.  In the case of
$k\notin\set{p,q}$, the permutation operator $\Pi_{pq}$ commutes with
$X_k$ and annihilates on $\ket\psi$, so
\begin{align}
  \sum_{k\in\ZZ_N} \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\notin\set{p,q}}}
  h_{pq} w_k X_k \ket\psi
  = E_0 \tilde X_{\m w} \ket\psi
  - \sum_{k\in\ZZ_N} \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\in\set{p,q}}}
  h_{pq} w_k X_k \ket\psi,
\end{align}
where $E_0 = \sum_{\set{p,q}\in\C_N\p{2}} h_{pq}$ is the interaction
energy of PS states $\ket\psi\in\M_0$.  Simplifying
\begin{align}
  \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\in\set{p,q}}} h_{pq}
  = \f12 \sum_{\substack{p,q\in\ZZ_N\\k\in\set{p,q}}} h_{pq}
  = \f12 \sum_{p\in\ZZ_N} h_{pk} + \f12 \sum_{q\in\ZZ_N} h_{kq}
  = h_k,
  &&
  h_k \equiv \sum_{p\in\ZZ_N} h_{pk},
\end{align}
we can write
\begin{align}
  \sum_{k\in\ZZ_N} \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\notin\set{p,q}}}
  h_{pq} w_k X_k \ket\psi
  = E_0 \tilde X_{\m w} \ket\psi
  - \sum_{k\in\ZZ_N} h_k w_k X_k \ket\psi.
\end{align}
The terms in \eqref{eq:single_body_diagnosis_start} with
$k\in\set{p,q}$, meanwhile, are
\begin{align}
  \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\in\set{p,q}}}
  h_{pq} w_k \Pi_{pq} X_k \ket\psi
  = \sum_{k,q\in\ZZ_N} h_{kq} w_k X_q \ket\psi,
\end{align}
so in total
\begin{align}
  H_0 X_{\m w} \ket\psi
  = E_0 X_{\m w} \ket\psi
  + \sum_{p\in\ZZ_N} \p{\sum_{q\in\ZZ_N} h_{pq} w_q - h_p w_p}
  X_p \ket\psi.
  \label{eq:single_body_diagnosis_end}
\end{align}
The action of the single-body perturbation $X_{\m w}$ on a
permutationally symmetric state therefore generates an eigenstate of
$H_0$ with energy $E_0+\Delta$ if the vector
$\v{\m w}\equiv\sum_{k\in\ZZ_N} w_k \ket{k}$ satisfies the eigenvalue
equation
\begin{align}
  \p{\m h - \diag\v h} \c \v{\m w} = \Delta \v{\m w},
  \label{eq:single_body_eig}
\end{align}
where $\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ is a matrix of
all couplings $h_{pq}$; the vector
$\v h \equiv \sum_{p\in\ZZ_N} h_p \ket{p} = \sum_{p,q\in\ZZ_N} h_{pq}
\ket{p}$ is the sum of all columns of $\m h$; and $\diag\v h$ is a
matrix with $\v h$ on the diagonal and zeroes everywhere else.

When the interaction Hamiltonian $H_0$ is translationally invariant,
the single-body eigenvalue problem in \eqref{eq:single_body_eig} is
solvable analytically.  In this case, the couplings $h_{pq}$ depend
only on the separation $\abs{p-q}$, so eigenvectors of $\m h$ are
therefore plane waves of the form
\begin{align}
  \v{\m w}_k \equiv \sum_{p\in\ZZ_L^D} e^{ip\c k} \ket{p},
\end{align}
where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
sites are indexed by vectors $p\in\ZZ_L^D$, and wavenumbers take on
values $k\in\ZZ_L^D\times2\pi/L$.  The corresponding eigenvalues of
$\m h$ can be determined by expanding
\begin{align}
  \m h\c{\m w}_k
  = \sum_{p,q\in\ZZ_L^D} h_{pq} e^{iq\c k} \ket{p}
  = \sum_{p,d\in\ZZ_L^D} h_{p,p+d} e^{i\p{p+d}\c k} \ket{p}
  = \sum_{d\in\ZZ_L^D} h_{0,d} \cos\p{d\c k} \v{\m w}_k,
\end{align}
where the imaginary terms vanish because $h_{0,d}=h_{0,-d}$.  The
remainder of \eqref{eq:single_body_eig} that we need to sort out is
$\diag\v h$, where all
$h_p=\sum_{q\in\ZZ_L^D}h_{pq}=\sum_{d\in\ZZ_L^D}h_{0,d}$ are equal,
which implies that $\diag\v h=\sum_{d\in\ZZ_L^D}h_{0,d}$ is a scalar.
We thus find that
\begin{align}
  \p{\m h - \diag\v h}\c\v{\m w}_k = \Delta_k \v{\m w}_k,
  &&
  \Delta_k \equiv \sum_{d\in\ZZ_L^D} h_{0,d} \sp{\cos\p{d\c k}-1}.
\end{align}

@@

@@

  H_0 \tilde X_{\m w} \ket\psi
  = \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}}}

@@

The sum in \eqref{eq:diagnosis_start} has terms with $p,q\notin k$,
terms with $p,q\in k$, and terms with only one of $p$ or $q\in k$.
The permutation operator $\Pi_{pq}$ acts trivially on $X_k\ket\psi$
when $p,q\notin k$, so

@@

  \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}\\p,q\notin k}}

@@

  = E_0 \tilde X_{\m w} \ket\psi
  - \sum_{k\in\D_N\p{M}} h^{\t{D}}_k w_k X_k \ket\psi,

@@

where $E_0 = \sum_{\set{p,q}\in\C_N\p{2}} h_{pq}$ is the interaction
energy of PS states $\ket\psi\in\M_0$, and

@@

  h^{\t{D}}_k \equiv
  \sum_{\substack{\set{p,q}\in\C_N\p{2}\\p\in k~\t{or}~q\in k}} h_{pq}
  = \sum_{p\in k} h_p - \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b},

@@

  h_p \equiv \sum_{q\in\ZZ_N} h_{pq}.

@@

Defining a linear operator $\m h_{\t{D}}$ on the space of
dimension-$M$ tensors, whose action on $\m w$ is a new dimension-$M$
tensor with entries

@@

  \m h_{\t{D}}\p{\m w}_k \equiv h^{\t{D}}_k w_k,
  \label{eq:multi_body_op_diag}

@@

we can write the terms in \eqref{eq:diagnosis_start} with
$p,q\notin k$ as
  \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}\\p,q\notin k}}
  = E_0 \tilde X_{\m w} \ket\psi
  - \sum_{k\in\D_N\p{M}} \m h_{\t{D}}\p{\m w}_k X_k \ket\psi.
The terms in \eqref{eq:diagnosis_start} with $p,q\in k$ are
  \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}\\p,q\in k}}
  = \sum_{k\in\D_N\p{M}} \sum_{\substack{\set{a,b}\in\C_M\p{2}}}
  h_{k_ak_b} w_{k_{a\lra b}} X_k \ket\psi,
where $k_{a\lra b}$ denotes a list that is equal to
$k=\p{k_1,k_2,\cdots,k_M}$ but with $k_a$ and $k_b$ swapped, and we
have relabeled $k_a$ and $k_b$ to change $w_k X_{k_{a\lra b}}$ to
$w_{k_{a\lra b}} X_k$.  To write down these terms succinctly, we
define a linear operator $\m h_{\t{P}}$ on the space of dimension-$M$
tensors, whose action on $\m w$ is a new dimension-$M$ tensor with
entries
  \m h_{\t{P}}\p{\m w}_k
  \equiv \sum_{\substack{\set{a,b}\in\C_M\p{2}}}
  h_{k_ak_b} w_{k_{a\lra b}},
  \label{eq:multi_body_op_perm}
such that
  \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}\\p,q\in k}}
  = \sum_{k\in\D_N\p{M}} \m h_{\t{P}}\p{\m w}_k X_k \ket\psi.
Finally, the terms in \eqref{eq:diagnosis_start} with only one of

@@

  \sum_{k\in\D_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  = \sum_{k\in\D_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \sum_{a\in\ZZ_M} h_{pk_a} w_{k_{a:p}} X_k \ket\psi,

@@

where $k_{a:p}$ denotes a list that is equal to $k$ but with $k_a$
replaced by $p$, and we have relabeled $k_a$ and $p$ to change
$w_k X_{k_{a:p}}$ to $w_{k_{a:p}} X_k$.  For each fixed $a$, the sum
over $p$ above is essentially a contraction of the matrix $\m h$ with
the $a$-th index of the tensor $\m w$.  Defining a linear operator
$\m h_a$ on the space of dimension-$M$ tensors, whose action on $\m w$
is a new dimension-$M$ tensor with entries

@@

  \m h_a\p{\m w}_k

@@

  h_{pk_a} w_{k_1 k_2 \cdots k_{a-1} p k_{a+1} \cdots k_M},
  \label{eq:multi_body_op_contract}

@@

we can expand
  \sum_{k\in\D_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  = \sum_{k\in\D_N\p{M}} \sum_{a\in\ZZ_M}
  \m h_a\p{\m w}_k X_k \ket\psi.
Altogether, we find that

@@

  = E_0 X_{\m w} \ket\psi + X_{\hat{\m h}\p{\m w}} \ket\psi,

@@

  \equiv \sum_{a\in\ZZ_M} \m h_a\p{\m w}
  + \m h_{\t{P}}\p{\m w} - \m h_{\t{D}}\p{\m w},

@@

is defined in terms of the tensors $\m h_a\p{\m w}$,
$\m h_{\t{P}}\p{\m w}$, and $\m h_{\t{D}}\p{\m w}$, whose entries are
respectively provided in \eqref{eq:multi_body_op_contract},
\eqref{eq:multi_body_op_perm}, and \eqref{eq:multi_body_op_diag}.  The
action of the $M$-body perturbation $X_{\m w}$ on a permutationally
symmetric state therefore generates an eigenstate of $H_0$ with energy
$E_0+\Delta$ if the coefficient tensor $\m w$ satisfies the
generalized eigenvalue equation $\hat{\m h}\p{\m w} = \Delta \m w$,
which can be written in the form

@@

  \v{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k \ket{k},

@@

  \hat{\m h} \equiv \sum_{k\in\D_N\p{M}} \ket{k}
  \sp{\sum_{a\in\ZZ_M} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
    h_{pk_a} \bra{k_{a:p}}
    + \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b} \bra{k_{a\lra b}}
    - h^{\t{D}}_k \bra{k}}.

@@

If $M=1$, then $h^{\t{D}}_k=h_k$, so

@@

  \hat{\m h} \stackrel{M=1}{=} \sum_{k\in\ZZ_N} \ket{k}
  \sp{\sum_{\substack{p\in\ZZ_N\\p\ne k}} h_{pk} \bra{p} - h_k \bra{k}}
  = \m h - \diag\v h,

@@

which is precisely what we found in \eqref{eq:single_body_eig}.  If
$M=2$, then $h_{pq}^{\t{D}}=h_p+h_q-h_{pq}$, so

@@

  \hat{\m h} \stackrel{M=2}{=}
  \sum_{\p{k,\ell}\in\D_N\p{2}} \ket{k,\ell}
  \sp{\sum_{\substack{p\in\ZZ_N\\p\notin\set{k,\ell}}}
    \p{ h_{pk} \bra{p,\ell} + h_{p\ell} \bra{k,p} }
    + h_{k\ell} \bra{\ell,k}
    - \p{h_k + h_\ell - h_{k\ell}} \bra{k,\ell}}.

@@

@@

Solving the eigenvalue problem in \eqref{eq:multi_body_eig_mat} to
find $M$-body perturbations $X_{\m w}$ that generate states of
definite interaction energy nominally requires diagonalizing the
matrix $\hat{\m h}$ with dimensions
$N!_M\times N!_M\sim N^M\times N^M$, where
$N!_M\equiv\prod_{j=0}^{M-1}\p{N-j}$ denotes a falling factorial.

If the interaction Hamiltonian $H_0$ has any spatial symmetries,
however, then we can use these symmetries to reduce the complexity of
this eigenvalue problem.  Translational invariance, for example, can
reduce this problem to that of diagonalizing a matrix with dimensions
$\le{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times N^{M-1}$,
and isotropy can further reduce the problem size by factors that
depend on the lattice geometry.  The complexity of diagonalizing a
matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a reduction
of $d$ by a factor of 2 yields practically significant gains (a factor
of $2^3=8$) in the runtime of diagonalization.  Here, we provide a
general prescription for exploiting the spatial symmetries of $H_0$ to
simplify the eigenvalue problem in \eqref{eq:multi_body_eig_mat}.

Our task is essentially to express the multi-body eigenvalue problem
in \eqref{eq:multi_body_eig_mat} in a basis that respects the spatial
symmetries of $H_0$.  These symmetries essentially break the set of
choices $\C_N\p{M}$ into equivalence classes $\S_M$\footnote{For
  brevity, we will suppress the explicit dependence of $\S_M$ on $N$,
  as well as the dependence of $\S_M$ on any other factors such as
  lattice geometry.}, where elements $s\in\S_M$ are subsets
$s\subset\C_N\p{M}$ that are invariant under the appropriate symmetry
group action.  In the case of a translationally invariant and
isotropic system with two-body perturbations, for example, each
equivalence class $s\in\S_2$ would consist of all pairs
$\set{p,q}\in\C_N\p{2}$ that are a fixed distance $\abs{p-q}=d_s$
apart, i.e.~$s=\set{\set{p,q}\in\C_N\p{2}:\abs{p-q}=d_s}$.
